Chr2 hotspot preliminary analyses

### 2019-07-01 analysis/paper/Chr2-prelim-fewer-figs.Rmd
started working with "...-fewer-figs.Rmd" file to speed up interactive work.

## Reading pleiotropy test results

@@

```{r, plotly_by_local, fig.height=ceiling(nlocal / 2)}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) + theme(legend.position = "none")
plotly::ggplotly(p)
```

```{r, plotly_by_nonlocal, fig.height=ceiling(nnonlocal / 4)}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ nonlocal_symbol, ncol = 4, nrow = ceiling(nnonlocal / 4)) + theme(legend.position = "none")
plotly::ggplotly(p)
```

@@

>>>>

## Reading pleiotropy test results (for local-nonlocal pairs)

@@

```{r, plotly_by_local, fig.height=ceiling(nlocal / 2), eval = FALSE}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) + theme(legend.position = "none")
plotly::ggplotly(p)
```

```{r, plotly_by_nonlocal, fig.height=ceiling(nnonlocal / 4), eval = FALSE}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ nonlocal_symbol, ncol = 4, nrow = ceiling(nnonlocal / 4)) + theme(legend.position = "none")
plotly::ggplotly(p)
```

@@

## Reading results for additional pleiotropy tests

Above we considered only those pleiotropy tests that paired a nonlocal trait with
a local trait. We also have results for the pairwise pleiotropy tests that involve pairs of 
local traits and pairs of nonlocal traits.

```{r read_pleio_results_local}
results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10 + 1)
dir(results_dir) -> fns
mylist_local <- list()
for (i in 1:length(fns)){
  mylist_local[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
  as_tibble() %>%
  mutate(filename = fns[i]) %>%
  mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
  filter(lod == max(lod), trace == "profile1") %>%
  select(gene1_id, gene2_id, lod, filename)
}
```

```{r read_pleio_results_nonlocal}
results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10)
dir(results_dir) -> fns
mylist_nonlocal <- list()
for (i in 1:length(fns)){
  mylist_nonlocal[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
  as_tibble() %>%
  mutate(filename = fns[i]) %>%
  mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], 
         gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
  filter(lod == max(lod), trace == "profile1") %>%
  select(gene1_id, gene2_id, lod, filename)
}
```

### Convert pleiotropy results lists to tibbles

```{r}
local_annot3 <- local_annot2 %>%
  select(annot.id, symbol)
local_pleio <- mylist_local %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(local_annot3, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(local_annot3, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol)

nonlocal_annot3 <- nonlocal_annot %>%
  select(annot.id, symbol)
nonlocal_pleio <- mylist_nonlocal %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(nonlocal_annot3, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(nonlocal_annot3, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol)

I now bind all three tibbles that contain pleiotropy results into a single tibble.

```{r}
pleio_results_mat <- foo %>%
  select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
  rename(gene1_id = local_gene_id, 
         gene2_id = nonlocal_gene_id, 
         gene1_symbol = local_symbol,
         gene2_symbol = nonlocal_symbol
         ) %>%
  bind_rows(local_pleio, nonlocal_pleio) %>%
  hotspots::tibble_to_matrix(symmetric = TRUE)
```

```{r}
pleio_results_mat %>%
  is.na() %>%
  sum()
```

```{r}
pleio_results_mat %>%
  dist() %>%
  hclust()
```

```{r}
foobar <- pleio_results_mat %>%
  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) %>%
  fastgreedy.community()
foobar$membership %>% table()
t1 <- tibble(gene = foobar$names, 
       community_membership = foobar$membership, 
       local = foobar$names %in% local_annot3$annot.id
       )
  table(t1$community_membership, t1$local)
```

## Clustering

```{r, hclust1}
pleio_results_mat %>%
  dist() %>%
  hclust() %>% 
  plot(cex = 0.1)
```

Check out https://genomicsclass.github.io/book/pages/clustering_and_heatmaps.html for helpful code on clustering gene expressions.

```{r rafalib_plot}
pleio_results_mat %>%
  dist() %>%
  hclust() %>% 
  rafalib::myplclust(
                     lab.col = 1 + (rownames(pleio_results_mat) %in% local_annot3$annot.id), cex = 0.1
                     )
```

```{r}
hclusters <- pleio_results_mat %>%
  dist() %>%
  hclust() %>% 
  cutree(h = 30)
table(true=rownames(pleio_results_mat) %in% local_annot3$annot.id, cluster=hclusters)
```

## Clustering of the rectangular nlocal by nnonlocal matrix

First, we make the matrix from the tibble `foo`.

```{r}
foomat <- foo %>%
  tibble_to_matrix()
```

```{r}
foomat_hc <- foomat %>%
  dist() %>%
  hclust
```

```{r}
plot(foomat_hc, cex = 0.7)
rafalib::myplclust(foomat_hc, cex = 0.7)
```

```{r}
groups <- foomat_hc %>%
  cutree(k = 6) %>%
  as_tibble(rownames = "gene_symbol") %>%
  rename(group_number = value) %>%
  arrange(desc(group_number))

## Heatmaps

```{r, load_colors}
library(RColorBrewer) 
library(genefilter)
library(gplots)
```

```{r define_color}
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
heatmap.2(foomat, 
          #labCol=tissue,
          trace="none", 
          #ColSideColors=cols, 
          col=hmcol,
          cexCol = 0.2, cexRow = 0.4
      )

## Mediation statistic clustering

```{r, loddiff_clust}
foo %>%
  select(1, 2, 4) %>% # choose lod_diff as the stat of interest
  tibble_to_matrix() %>%
  heatmap.2( 
          #labCol=tissue,
          trace="none", 
          #ColSideColors=cols, 
          col=hmcol,
          cexCol = 0.2, 
          cexRow = 0.4
      )

```{r, loddiffprop_clust}
foo %>%
  select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
  tibble_to_matrix() %>%
  heatmap.2( 
          #labCol=tissue,
          trace="none", 
          #ColSideColors=cols, 
          col=hmcol,
          cexCol = 0.2, cexRow = 0.4
      )

```{r load_d3heat}
library(d3heatmap)
```

### Binarizing mediation matrix before clustering

```{r threshold_clust}
med_thresh <- 2
foomat <- foo %>%
  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
  tibble_to_matrix()
  apply(FUN = function(x) x <= med_thresh, X = foomat, MARGIN = 2) %>%
    d3heatmap()
```

```{r}
foo %>%
  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
  tibble_to_matrix() %>%
  d3heatmap(cexRow = 0.2, cexCol = 0.2)
```

## IDEAS

Maybe we think about clustering of subsets. So, subset on, say, pleiotropy values then cluster the resulting subsets separately??

Also, think about writing a procedure for identifying "pleiotropy groups".

Also, we might first remove those nonlocal traits that are clearly mediated by, eg, Hnf4a. Can we then use pleiotropy tests to inform what remains?

## kmeans

```{r kmeans}
set.seed(01-28-1978)
# replace NA with zeros
pleio_resd0 <- apply(FUN = function(x){x[is.na(x)] <- 0; return(x)}, X = pleio_results_mat, MARGIN = 2)
#km <- kmeans(t(e[1:2,]), centers=7)
#names(km)
